Abstract

### 2025-03-02 output/Main_text.Rmd
Correct reference to scale invariance
"Norm invariance" was incorrectly referred to as "scale invariance", making the discourse contradictory.
(Also, a minor reference to scale invariance in the "Covariance-based version" section is improved)

The second property derives straightforwardly from
Equation \@ref(eq:distance-param),
and the third one is implicit in the definition of the $`r MDISC_SYM`$.
However, when an item measures only dimension $`r DIM_INDEX`$,
simple arithmetic can show that $`r MDISC_UNIDIM_GENERALIZED_EQ`$.
In the covariance-based case, this means that $`r MDISC_COV_BASED_UNIDIM_EQ`$
and, as $`r COV_SYM_DIAG`$ is generally different from 1,
$`r MDISC_COV_PARAM`$ does not fulfill the first property
(we shall refer to this property as *scale invariance*, from now on).
Nevertheless, it can be interpreted as a *scaled version* of the parameter,
with the standard deviation of the corresponding dimension as scaling factor.
This transformation may be made in the unidimensional case as well,
thus finding a *standard-metric* discrimination parameter
(i.e., referred to a unitary variance space).
Moreover, it is important to notice that,
when the latent trait inner product Gram matrix
is defined as $`r INNER_PROD_COV_COND`$,
the norm of a trait vector in the latent space is given by
the Mahalanobis distance, i.e., $`r MAH_DIST_EQ`$ [@mahalanobis_reprint_2018].

@@

With that purpose in mind, we have defined two new versions of these parameters:
One that takes into account the whole covariance structure,
and another one that only considers the correlations among latent dimensions.
The covariance-based parameters have the drawback of not fulfilling
the first of Reckase & McKinley's [-@reckase_discriminating_1991] properties,
namely, the scale invariance property.
However, we have provided a rationale for the violation of this property,
arguing that it is also paralleled in unidimensional IRT.
Given its equivalence to the inner product definition in a space with
an identity covariance matrix,
it also refers the item parameters to an orthonormal metric,
comparable among different latent spaces.
More importantly, the covariance-based version
of the multidimensional parameters is invariant to
changes of latent-space basis.
Once again, this is relevant in applications where
potentially different bases are involved.
Furthermore, we have found that using this version yields a norm in
the latent space given by the Mahalanobis distance.
This statistical distance not only takes
the latent covariance structure into account,
but also makes the latent space norm invariant to changes of basis.
Given these considerations,
we propose adopting the covariance-based version of
the multidimensional parameters as the most general definition,
and refer to them simply as
the $`r MDISC_SYM`$ and $`r MIL_PARAM`$ parameters.
The correlation-based version, on the other hand,
has the advantage of being scale-invariant,
which is convenient for fidelity when representing the items.
However, as it has been defined ad-hoc to fulfill this specific property,
it lacks the other desirable ones the covariance-based version has
(i.e., scale invariance, and yielding a diagonal latent covariance matrix in
the corresponding orthogonal basis).
Therefore, we recommend its use for graphical representation purposes only,
although in many applications both versions will be coincident, fortunately.
As for the agnostic version, we recommend it only in
the (unlikely) case of having no information about the latent space structure,
or when such structure is irrelevant
(e.g., plotting the discrimination parameters for interpretation purposes).

>>>>

The second property derives straightforwardly from
Equation \@ref(eq:distance-param),
and the third one is implicit in the definition of the $`r MDISC_SYM`$.
However, when an item measures only dimension $`r DIM_INDEX`$,
simple arithmetic can show that $`r MDISC_UNIDIM_GENERALIZED_EQ`$.
In the covariance-based case, this means that $`r MDISC_COV_BASED_UNIDIM_EQ`$
and, as $`r COV_SYM_DIAG`$ is generally different from 1,
$`r MDISC_COV_PARAM`$ does not fulfill the first property
(we shall refer to this property as *scale invariance*, henceforth).
Nevertheless, $`r MDISC_COV_PARAM`$ can be interpreted as
a *scaled version* of the parameter,
with the standard deviation of the corresponding dimension as scaling factor.
This transformation may be made in the unidimensional case as well,
thus finding a *standard-metric* discrimination parameter
(i.e., referred to a unitary variance space).
Moreover, it is important to notice that,
when the latent trait inner product Gram matrix
is defined as $`r INNER_PROD_COV_COND`$,
the norm of a trait vector in the latent space is given by
the Mahalanobis distance, i.e., $`r MAH_DIST_EQ`$ [@mahalanobis_reprint_2018].

@@

With that purpose in mind, we have defined two new versions of these parameters:
One that takes into account the whole covariance structure,
and another one that only considers the correlations among latent dimensions.
The covariance-based parameters have the drawback of not fulfilling
the first of Reckase & McKinley's [-@reckase_discriminating_1991] properties,
namely, the scale invariance property.
However, we have provided a rationale for the violation of this property,
arguing that it is also paralleled in unidimensional IRT.
Given its equivalence to the inner product definition in a space with
an identity covariance matrix,
it also refers the item parameters to an orthonormal metric,
comparable among different latent spaces.
More importantly, the covariance-based version
of the multidimensional parameters is invariant to
changes of latent-space basis.
Once again, this is relevant in applications where
potentially different bases are involved.
Furthermore, we have found that using this version yields a norm in
the latent space given by the Mahalanobis distance.
This statistical distance not only takes
the latent covariance structure into account,
but also makes the latent space norm invariant to changes of basis.
Given these considerations,
we propose adopting the covariance-based version of
the multidimensional parameters as the most general definition,
and refer to them simply as
the $`r MDISC_SYM`$ and $`r MIL_PARAM`$ parameters.
The correlation-based version, on the other hand,
has the advantage of being scale-invariant,
which is convenient for fidelity when representing the items.
However, as it has been defined ad-hoc to fulfill this specific property,
it lacks the other desirable ones the covariance-based version has
(i.e., norm invariance, and yielding a diagonal latent covariance matrix in
the corresponding orthogonal basis).
Therefore, we recommend its use for graphical representation purposes only,
although in many applications both versions will be coincident, fortunately.
As for the agnostic version, we recommend it only in
the (unlikely) case of having no information about the latent space structure,
or when such structure is irrelevant
(e.g., plotting the discrimination parameters for interpretation purposes).
